On the use of decision trees to identify the local extrema in parallel global optimization algorithms

### 2023-01-30 2023/PAVT/ParallelDecisionTree.tex
proofread the last section

% переводить до этого раздела
\section{Результаты экспериментов}\label{SecR}

@@

To demonstrate the efficiency of  investigated global search algorithm, here we present the results of  comparing this one with well known methods DIRECT and DIRECTl. Table \ref{tab:1} presents the averaged numbers of iterations executed by respective methods when  solving a series of GKLS problems. The sign «>» points to the situation when not all problems are  solved. In these cases, the numbers of non-solved problems are given in brackets. As one can see,  GSA overcomes DIRECT and DIRECTl methods in average number of iterations necessary for  solving the problems with the same precision.

@@

Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (GSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented. The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard) of dimensionalities 2, 3, 4, and 5 from \cite{fio_bib19}. The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.  In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of speedup in time relative to the sequential run are compared. The  parallelization was performed using MPI technology, the run was performed on 8 processes.

@@

		N & Problem class & \multicolumn{2}{c|}{Averaged number of iterations} & \multicolumn{2}{c|}{Speedup} \\ \hline
		& ~ & GSAA & Decision tree & GSAA & Decision tree \\ \hline
		& Simple & 3823,1   & 485,6  & 7,4   & 12,7 \\ \hline
		2  & Hard & 787,6    & 304,2  & 13,8  & 14,9 \\ \hline
		& Simple & 5700,5   & 720,1  & 5,6   & 14,2  \\ \hline
		3  & Hard & 2411,9   & 634,3  & 5,0   & 5,9  \\ \hline
		& Simple & 31101,5  & 1037,4 & 5,1   & 7,1  \\ \hline
		4  & Hard & 10418,1  & 816,2  & 6,9   & 5,0  \\ \hline
		& Simple & 163313,5 & 1204,2 & 4,8   & 4,1 \\ \hline
		5  & Hard & 27524,5  & 999,4  & 4,6   & 6,3  \\ \hline
	\end{tabular}
\end{table}

@@

So far, as a result of the work done, we succeeded to combine the global search algorithm with a local  optimization method.  Unlike the known multistart schemes, the decision to run the local method is made using the decision  tree. The use of such a combination of methods allows accelerating the work of algorithm considerably.

The parallel version of the algorithm preserves the properties of its sequential prototype that was  confirmed by he numerical experiments on solving a series of several hundred problems of various  dimensionalities. The scheme we proposed allows utilizing the advantages of both approaches – of the  parallelization as well as of fast search of local extrema.

>>>>

% переводить до этого раздела
\section{Results and Conclusion}\label{SecR}

@@

To demonstrate the efficiency of  investigated global search algorithm, here we present the results of  comparing this one with well known methods DIRECT and DIRECTl. Table \ref{tab:1} presents the averaged numbers of iterations executed by respective methods when  solving a series of GKLS problems. The sign ``>'' points to the situation when not all problems are  solved. In this case, the number of non-solved problems is given in brackets. As one can see,  GSA overcomes DIRECT and DIRECTl methods in average number of iterations necessary for solving the problems with the same precision.

@@

Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (AGSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented. The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{fio_bib19}) of dimensionalities 2, 3, 4, and 5. The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.  In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of time speedup relative to the sequential run are compared. The parallelization was performed using MPI technology, the run was performed on 8 processes.

@@

		N & Problem class & \multicolumn{2}{c|}{Averaged number of iterations} & \multicolumn{2}{c|}{Speedup} \\ \hline
		& ~ & AGSA & Decision tree & AGSA & Decision tree \\ \hline
		& Simple & 3823.1   & 485.6  & 7.4   & 12.7 \\ \hline
		2  & Hard & 787.6    & 304.2  & 13.8  & 14.9 \\ \hline
		& Simple & 5700.5   & 720.1  & 5.6   & 14.2  \\ \hline
		3  & Hard & 2411.9   & 634.3  & 5.0   & 5.9  \\ \hline
		& Simple & 31101.5  & 1037.4 & 5.1   & 7.1  \\ \hline
		4  & Hard & 10418.1  & 816.2  & 6.9   & 5.0  \\ \hline
		& Simple & 163313.5 & 1204.2 & 4.8   & 4.1 \\ \hline
		5  & Hard & 27524.5  & 999.4  & 4.6   & 6.3  \\ \hline
	\end{tabular}
\end{table}

@@

So far, as a result of the work done, we succeeded to combine the global search algorithm with a local  optimization method.  Unlike the known multistart schemes, the decision to run the local method is made using the decision  tree. The use of such a combination of methods allows considerably accelerating the algorithm.

The parallel version of the algorithm preserves the properties of its sequential prototype that was  confirmed by he numerical experiments on solving a series of several hundred problems of various  dimensionalities. The proposed scheme allows us utilizing the advantages of both parallelization as well as fast search of local extrema.
